---
title: "Glasgow TB ACF"
output: html_notebook
---

### 1. Libraries and functions

#### 1.1 Libraries

Load the required libraries.

```{r, message=F, warning=F}
library(tidyverse)
library(sf)
library(here)
library(readxl)
library(scales)
library(DT)
library(brms)
library(tidybayes)
library(patchwork)
library(marginaleffects)
library(ggrepel)
library(scico)
library(ggdensity)
library(ggpubr)
library(ggsn)

```

#### 1.2 Helper functions

Functions that we will use throughout the script

```{r}
#labeller for years
year_labels <- c(1950:1963)

#The Glasgow mass minuture chest X-ray campaign happened between 11th March and 12th April 1957
#Segment for graphs to match ACF period
acf_start <- decimal_date(ymd("1957-03-11"))
acf_end <- decimal_date(ymd("1957-04-12"))


```

Function for counterfactual plots

```{r}

plot_counterfactual <- function(model_data, model, population_denominator, outcome, grouping_var=NULL, ...){
  
  #labeller for years
  year_labels <- c(1950:1963)

  #The Glasgow mass minuture chest X-ray campaign happened between 11th March and 12th April 1957
  #Segment for graphs to match ACF period
  acf_start <- decimal_date(ymd("1957-03-11"))
  acf_end <- decimal_date(ymd("1957-04-12"))

  summary <- {{model_data}} %>%
    select(year, year2, y_num, acf_period, {{population_denominator}}, {{outcome}}, {{grouping_var}}) %>%
    add_epred_draws({{model}}) %>%
    group_by(year2, acf_period, {{grouping_var}}) %>%
    median_qi() %>%
    mutate(.epred_inc = .epred/{{population_denominator}}*100000,
          .epred_inc.lower = .epred.lower/{{population_denominator}}*100000,
          .epred_inc.upper = .epred.upper/{{population_denominator}}*100000) %>%
    mutate(acf_period = case_when(acf_period=="a. pre-acf" ~ "Before Intervention",
                                  acf_period=="c. post-acf" ~ "Post Intervention"))



  #create the counterfactual (no intervention), and summarise
  
  counterfact <-
    add_epred_draws(object = {{model}},
                    newdata = {{model_data}} %>%
                                  select(year, year2, y_num, {{population_denominator}}, {{grouping_var}}, {{outcome}}) %>%
                                  mutate(acf_period = "a. pre-acf")) %>%
    group_by(year2, acf_period, {{grouping_var}}) %>%
    median_qi() %>%
    mutate(.epred_inc = .epred/{{population_denominator}}*100000,
         .epred_inc.lower = .epred.lower/{{population_denominator}}*100000,
         .epred_inc.upper = .epred.upper/{{population_denominator}}*100000) %>%
    mutate(acf_period = case_when(acf_period=="a. pre-acf" ~ "Before Intervention",
                                acf_period=="c. post-acf" ~ "Post Intervention"))
  


  #plot the intervention effect
p <- summary %>%
    droplevels() %>%
    ggplot() +
    geom_ribbon(aes(ymin=.epred_inc.lower, ymax=.epred_inc.upper, x=year2, group = acf_period, fill=acf_period), alpha=0.5) +
    geom_ribbon(data = counterfact %>% filter(year>=1956), 
                aes(ymin=.epred_inc.lower, ymax=.epred_inc.upper, x=year2, fill="Counterfactual"), alpha=0.5) +
    geom_line(data = counterfact %>% filter(year>=1956), 
              aes(y=.epred_inc, x=year2, colour="Counterfactual")) +
    geom_line(aes(y=.epred_inc, x=year2, group=acf_period,  colour=acf_period)) +
    geom_point(data = {{model_data}}, aes(y={{outcome}}, x=year2, shape=acf_period), size=2) +
    geom_vline(aes(xintercept=acf_end), linetype=3) +
    theme_ggdist() +
    scale_y_continuous(labels=comma) +
    scale_x_continuous(labels = year_labels,
                       breaks = year_labels) +
    scale_fill_manual(values = c("#DE0D92", "grey50", "#4D6CFA") , name="") +
    scale_colour_manual(values = c("#DE0D92", "grey50", "#4D6CFA") , name="") +
    scale_shape_discrete(name="") +
    labs(
      x = "Year",
      y = "Case notification rate (per 100,000)",
      caption = "Mass miniature X-ray campaign period between dashed lines (11th March-12th April 1957)"
    ) +
    theme(legend.position = "bottom",
          panel.border = element_rect(colour = "grey78", fill=NA),
          title = element_text(size=14),
          axis.text.x = element_text(size=10, angle = 90, hjust=1, vjust=0.5),
          legend.text = element_text(size=12)) +
    guides(shape="none")

    facet_vars <- vars(...)

  if (length(facet_vars) != 0) {
    p <- p + facet_wrap(facet_vars)
  }
  p

}

```

Function for calculating  measures of change over time

```{r}

summarise_change <- function(model_data, model, population_denominator, grouping_var=NULL){

  #a. immediate change
  nd_immediate <- {{model_data}} %>%
    filter(year %in% c(1956:1957)) %>%
    select(acf_period, year, y_num, {{population_denominator}}, {{grouping_var}})

  #Calcuate incidence per draw, then summarise.
  immediate_change <- add_epred_draws({{model}},
                                      newdata=nd_immediate) %>%
    mutate(epred_inc100k = .epred/{{population_denominator}}) %>%
    group_by(.draw, {{grouping_var}}) %>%
    mutate(acf_inc100k_diff = last(epred_inc100k)-first(epred_inc100k),
           acf_inc100k_rr = last(epred_inc100k)/first(epred_inc100k)) %>%
    ungroup() %>%
    group_by({{grouping_var}}) %>%
    mean_qi(acf_inc100k_diff, acf_inc100k_rr) %>%
    mutate(change = "Immediate change") %>%
    ungroup()
  
  #b. post-ACF change
  nd_post <- {{model_data}} %>%
    filter(year %in% c(1956,1958)) %>%
    select(acf_period, year, y_num, {{population_denominator}}, {{grouping_var}})

  #Calcuate incidence per draw, then summarise.
  post_change <- add_epred_draws({{model}},
                                      newdata=nd_post) %>%
    mutate(epred_inc100k = .epred/{{population_denominator}}) %>%
    group_by(.draw, {{grouping_var}}) %>%
    mutate(acf_inc100k_diff = last(epred_inc100k)-first(epred_inc100k),
           acf_inc100k_rr = last(epred_inc100k)/first(epred_inc100k)) %>%
    ungroup() %>%
    group_by({{grouping_var}}) %>%
    mean_qi(acf_inc100k_diff, acf_inc100k_rr) %>%
    mutate(change = "Post-ACF change") %>%
    ungroup()
  
  #c. change in slope post vs. pre-ACF
  slope_change <- {{model_data}} %>%
    select(year, year2, y_num, acf_period, {{population_denominator}}, {{grouping_var}}) %>%
    filter(year!=1957) %>%
    add_epred_draws({{model}}) %>%
    mutate(inc_100k = .epred/{{population_denominator}}*100000) %>%
    group_by(year, {{grouping_var}}, acf_period, ) %>%
    mean_qi(inc_100k) %>%
    ungroup() %>%
    mutate(n_years = length(year), .by=c(acf_period, {{grouping_var}})) %>%
    summarise(pct_change_epred_overall = (((last(inc_100k) - first(inc_100k))/first(inc_100k))),
              pct_change_lower_overall = (((last(.lower) - first(.lower))/first(.lower))),
              pct_change_upper_overall = (((last(.upper) - first(.upper))/first(.upper))),
      
              pct_change_epred_annual = (((last(inc_100k) - first(inc_100k))/first(inc_100k))/n_years),
              pct_change_lower_annual = (((last(.lower) - first(.lower))/first(.lower))/n_years),
              pct_change_upper_annual = (((last(.upper) - first(.upper))/first(.upper))/n_years),
              .by = c(acf_period, {{grouping_var}})) %>%
    distinct() %>%
    mutate(change = "Slope change")

  lst(immediate_change, post_change, slope_change)
    
}

```


Function for calculating difference from counterfactual

```{r}
calcuate_counterfactual <- function(model_data, model, population_denominator, grouping_var=NULL){
  
  #effect vs. counterfactual
  counterfact <-
      add_epred_draws(object = {{model}},
                      newdata = {{model_data}} %>%
                                    select(year, year2, y_num, {{population_denominator}}, {{grouping_var}}) %>%
                                    mutate(acf_period = "a. pre-acf")) %>%
      group_by(.draw, year, {{grouping_var}}, acf_period) %>%
      mutate(.epred_inc_counterf = .epred/{{population_denominator}}*100000, .epred_counterf=.epred)  %>%
      filter(year>1957) %>%
      ungroup() %>%
      select(year, {{population_denominator}}, .draw, .epred_counterf, .epred_inc_counterf)
  
  #Calcuate incidence per draw, then summarise.
  post_change <-
      add_epred_draws(object = {{model}},
                      newdata = {{model_data}} %>%
                                    select(year, year2, y_num, {{population_denominator}}, {{grouping_var}}, acf_period)) %>%
      group_by(.draw, year, {{grouping_var}}, acf_period) %>%
      mutate(.epred_inc = .epred/{{population_denominator}}*100000)  %>%
      filter(year>1957) %>%
      ungroup() %>%
      select(year, {{population_denominator}}, {{grouping_var}}, .draw, .epred, .epred_inc) 
  
  #for the overall period
    counterfact_overall <-
      add_epred_draws(object = {{model}},
                      newdata = {{model_data}} %>%
                                    select(year, year2, y_num, {{population_denominator}}, {{grouping_var}}) %>%
                                    mutate(acf_period = "a. pre-acf")) %>%
      group_by(.draw, {{grouping_var}}) %>%
      filter(year>1957) %>%
      ungroup() %>%
      select({{population_denominator}}, .draw, .epred)  %>%
      group_by(.draw) %>%
      summarise(.epred_counterf = sum(.epred)) %>%
      mutate(year = "Overall (1958-1963)")
  
  #Calcuate incidence per draw, then summarise.
  post_change_overall <-
      add_epred_draws(object = {{model}},
                      newdata = {{model_data}} %>%
                                    select(year, year2, y_num, {{population_denominator}}, {{grouping_var}}, acf_period)) %>%
      group_by(.draw, {{grouping_var}}) %>%
      filter(year>1957) %>%
      ungroup() %>%
      select({{population_denominator}}, {{grouping_var}}, .draw, .epred) %>%
      group_by(.draw, {{grouping_var}}) %>%
      summarise(.epred = sum(.epred)) 
  
  
counter_post <-
  left_join(counterfact, post_change) %>%
    mutate(cases_averted = .epred_counterf-.epred,
           pct_change = (.epred - .epred_counterf)/.epred_counterf,
           diff_inc100k = .epred_inc - .epred_inc_counterf,
           rr_inc100k = .epred_inc/.epred_inc_counterf) %>%
    group_by(year, {{grouping_var}}) %>%
    mean_qi(cases_averted, pct_change, diff_inc100k, rr_inc100k) %>%
    ungroup()

counter_post_overall <-
  left_join(counterfact_overall, post_change_overall) %>%
    mutate(cases_averted = .epred_counterf-.epred,
           pct_change = (.epred - .epred_counterf)/.epred_counterf) %>%
    group_by({{grouping_var}}) %>%
    mean_qi(cases_averted, pct_change) %>%
    ungroup() %>%
    mutate(year = "Overall (1958-1963)") 

lst(counter_post, counter_post_overall)

}



```

Function for tidying up counterfactuals (mostly for making nice tables)

```{r}

tidy_counterfactuals <- function(data){
  data %>%
  mutate(across(c(cases_averted:cases_averted.upper, diff_inc100k:diff_inc100k.upper), number_format(accuracy = 0.1, big.mark = ","))) %>%
  mutate(across(c(rr_inc100k:rr_inc100k.upper), number_format(accuracy = 0.01))) %>%
  mutate(across(c(pct_change:pct_change.upper), percent, accuracy=0.1)) %>%
  mutate(year = as.character(year),
            cases_averted = glue::glue("{cases_averted} ({cases_averted.lower} to {cases_averted.upper})"),
            pct_change = glue::glue("{pct_change} ({pct_change.lower} to {pct_change.upper})"),
            diff_inc = glue::glue("{diff_inc100k} ({diff_inc100k.lower} to {diff_inc100k.upper})"),
            rr_inc = glue::glue("{rr_inc100k} ({rr_inc100k.lower} to {rr_inc100k.upper})"))
}


tidy_counterfactuals_overall <- function(data){
  data %>%
  mutate(across(c(cases_averted:cases_averted.upper), number_format(accuracy = 0.1, big.mark = ","))) %>%
  mutate(across(c(pct_change:pct_change.upper), percent, accuracy=0.1)) %>%
  mutate(year = as.character(year),
            cases_averted = glue::glue("{cases_averted} ({cases_averted.lower} to {cases_averted.upper})"),
            pct_change = glue::glue("{pct_change} ({pct_change.lower} to {pct_change.upper})"))
}

```



### 2. Data

Import datasets for analysis

#### 2.1 Jonathan Golub's data

Import data from Jonathan Golub's paper (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4472641/), and summarise in a figure

```{r}

golub <- read_xlsx("2024_01_10_golub.xlsx")

golub_cxr <- golub %>%
  filter(!is.na(mass_cxr)) %>%
  separate(year_country, into = c("year", "country")) %>%
  mutate(year = as.numeric(year)) %>%
  filter(year<1980) %>%
  mutate(target_population = str_replace_all(sample, " ", ""),
         target_population = str_extract(target_population, "\\d+"))


```




#### 2.2 Shapefiles

Make a map of Glasgow wards

```{r}

glasgow_wards_1951 <- st_read(here("mapping/glasgow_wards_1951.geojson"))

```

```{r}

#read in Scotland boundary
scotland <- st_read(here("mapping/Scotland_boundary/Scotland boundary.shp"))



#make a bounding box for Glasgow
bbox <- st_bbox(glasgow_wards_1951) |> st_as_sfc()

#plot scotland with a bounding box around the City of Glasgow
scotland_with_bbox <- ggplot() +
  geom_sf(data = scotland, fill="antiquewhite") +
  geom_sf(data = bbox, colour = "#C60C30", fill="antiquewhite") +
  theme_void() +
  theme(panel.border = element_rect(colour = "grey78", fill=NA, linewidth = 0.5),
        panel.background = element_rect(fill = "#EAF7FA", size = 0.3))


#plot the wards
#note we tidy up some names to fit on map
glasgow_ward_map <- glasgow_wards_1951 %>%
  mutate(ward = case_when(ward=="Shettleston and Tollcross" ~ "Shettleston and\nTollcross",
                          ward=="Partick (West)" ~ "Partick\n(West)",
                          ward=="Partick (East)" ~ "Partick\n(East)",
                          ward=="North Kelvin" ~ "North\nKelvin",
                          ward=="Kinning Park" ~ "Kinning\nPark",
                          TRUE ~ ward)) %>%
  
  ggplot() +
  geom_sf(aes(fill=division)) +
  geom_sf_label(aes(label = ward), size=3, fill=NA, label.size = NA, colour="black", family = "Segoe UI") +
  #scale_colour_identity() +
  scale_fill_brewer(palette = "Set3", name="City of Glasgow Division") +
  theme_grey(base_family = "Segoe UI") +
  labs(x="",
       y="",
       fill="Division") +
  theme(legend.position = "top",
        
        panel.border = element_rect(colour = "grey78", fill=NA, linewidth = 0.5),
        panel.background = element_rect(fill = "antiquewhite", size = 0.3),
        panel.grid.major = element_line(color = "grey78")) +
  guides(fill=guide_legend(title.position = "top", title.hjust = 0.5, title.theme = element_text(face="bold"))) +
  scalebar(glasgow_wards_1951, dist = 2, dist_unit = "km",
             transform = TRUE, model = "WGS84", location="bottomleft")

#add the map of scotland as an inset
glasgow_ward_map + inset_element(scotland_with_bbox, 0.75, 0, 1, 0.4)

ggsave(here("figures/s1.png"), height=10, width = 12)


```



### 3. Denominators

Load in the datasets for denonomiators, and check for consistency.

```{r}

overall_pops <- read_xlsx(path = "2023-11-28_glasgow-acf.xlsx", sheet = "overall_population")

overall_pops %>%
  mutate(across(where(is.numeric) & !(year),  ~comma(.))) %>%
  datatable()

#shift year to midpoint
overall_pops <- overall_pops %>%
  mutate(year2 = year+0.5)

```

Note, we have three population estimates:

1. Population without institutionalised people or people in shipping
2. Population in institutions
3. Population in shipping

(Population in shipping is estimated from the 1951 census, so is the same for most years)

#### 3.1 Overall population

First, plot the total population

```{r}

overall_pops %>%
  ggplot() +
  geom_area(aes(y=total_population, x=year2), alpha=0.5, colour = "mediumseagreen", fill="mediumseagreen") +
  geom_point(aes(y=total_population, x=year2), colour = "mediumseagreen") +
  geom_vline(aes(xintercept=acf_start), linetype=3) +
  geom_vline(aes(xintercept=acf_end), linetype=3) +
  scale_y_continuous(labels=comma) +
  scale_x_continuous(labels = year_labels,
                     breaks = year_labels) +
  labs(
    title = "Glasgow Corporation: total population",
    subtitle = "1950 to 1963",
    x = "Year",
    y = "Population",
    caption = "Mid-year estimates\nMass miniature X-ray campaign period between dashed lines (11th March-12th April 1957)"
  ) +
  theme_ggdist()


```

Now the population excluding institutionalised and shipping population

```{r}

overall_pops %>%
  ggplot() +
  geom_area(aes(y=population_without_inst_ship, x=year2), alpha=0.5, colour = "purple", fill="purple") +
  geom_point(aes(y=population_without_inst_ship, x=year2), colour = "purple") +
  geom_vline(aes(xintercept=acf_start), linetype=3) +
  geom_vline(aes(xintercept=acf_end), linetype=3) +
  scale_y_continuous(labels=comma) +
  scale_x_continuous(labels = year_labels,
                     breaks = year_labels) +
  labs(
    title = "Glasgow Corporation: population excluding institutionalised and shipping",
    subtitle = "1950 to 1963",
    x = "Year",
    y = "Population",
    caption = "Mid-year estimates\nMass miniature X-ray campaign period between dashed lines (11th March-12th April 1957)"
  ) +
  theme_ggdist()


```

#### 3.2 Population by Ward

There are 5 Divisions containing 37 Wards in the Glasgow Corporation, with consistent boundaries over time.

```{r}
#look-up table for divisions and wards
ward_lookup <- read_xlsx(path = "2023-11-28_glasgow-acf.xlsx", sheet = "divisions_wards")


ward_pops <- read_xlsx(path = "2023-11-28_glasgow-acf.xlsx", sheet = "ward_population")

ward_pops %>%
  mutate(across(where(is.numeric) & !(year),  ~comma(.))) %>%
  datatable()

#shift year to midpoint
ward_pops <- ward_pops %>%
  mutate(year2 = year+0.5)

#Get the Division population
division_pops <- ward_pops %>%
  group_by(division, year) %>%
  summarise(population_without_inst_ship = sum(population_without_inst_ship, na.rm = TRUE),
            institutions = sum(institutions, na.rm = TRUE),
            shipping = sum(shipping, na.rm = TRUE),
            total_population = sum(total_population, na.rm = TRUE))

division_pops %>%
  mutate(across(where(is.numeric) & !(year),  ~comma(.))) %>%
  datatable()

```

Plot the overall population by Division and Ward

```{r}

division_pops %>%
  mutate(year2 = year+0.5) %>%
  ggplot() +
  geom_area(aes(y=total_population, x=year2, colour=division, fill=division), alpha=0.8) +
  geom_point(aes(y=total_population, x=year2, colour=division)) +
  geom_vline(aes(xintercept=acf_start), linetype=3) +
  geom_vline(aes(xintercept=acf_end), linetype=3) +
  scale_y_continuous(labels=comma) +
  facet_wrap(division~.) +
  scale_x_continuous(labels = year_labels,
                     breaks = year_labels,
                     guide = guide_axis(angle = 90)) +
  scale_fill_brewer(palette = "Set3", name = "") +
  scale_colour_brewer(palette = "Set3", name = "") +
  labs(
    title = "Glasgow Corporation: total population by Division",
    subtitle = "1950 to 1963",
    x = "Year",
    y = "Population",
    caption = "Mid-year estimates\nMass miniature X-ray campaign period between dashed lines (11th March-12th April 1957)"
  ) +
  theme_ggdist() +
  theme(legend.position = "bottom")


```

```{r}

ward_pops %>%
  ggplot() +
  geom_area(aes(y=total_population, x=year2, colour=division, fill=division), alpha=0.8) +
  geom_point(aes(y=total_population, x=year2, colour=division)) +
  geom_vline(aes(xintercept=acf_start), linetype=3) +
  geom_vline(aes(xintercept=acf_end), linetype=3) +
  scale_y_continuous(labels=comma) +
  facet_wrap(ward~., ncol=6) +
  scale_x_continuous(labels = year_labels,
                     breaks = year_labels,
                     guide = guide_axis(angle = 90)) +
  scale_fill_brewer(palette = "Set3", name="Division") +
  scale_colour_brewer(palette = "Set3", name = "Division") +
  labs(
    title = "Glasgow City: total population by Ward",
    subtitle = "1950 to 1963",
    x = "Year",
    y = "Population",
    caption = "Mid-year estimates\nMass miniature X-ray campaign period between dashed lines (11th March-12th April 1957)"
  ) +
  theme_ggdist() +
  theme(legend.position = "bottom")

ggsave(here("figures/s2.png"), height=10, width=12)

```

Approximately, how many person-years of follow-up do we have?

```{r}

overall_pops %>%
  ungroup() %>%
  summarise(across(year, length, .names = "years"),
            across(c(population_without_inst_ship, total_population), sum)) %>%
  mutate(across(where(is.double), comma)) %>%
  datatable()


```

Change in population by ward

```{r}

ward_pops %>%
  group_by(ward) %>%
  summarise(pct_change_pop = (last(population_without_inst_ship) - first(population_without_inst_ship))/first(population_without_inst_ship)) %>%
  mutate(pct_change_pop = percent(pct_change_pop)) %>%
  arrange(pct_change_pop) %>%
  datatable()
  


```


#### 3.3 Population by age and sex

```{r}

age_sex <- read_xlsx(path = "2023-11-28_glasgow-acf.xlsx", sheet = "age_sex_population") %>%
  pivot_longer(cols = c(male, female),
               names_to = "sex")

#collapse down to smaller age groups to be manageable
age_sex <- age_sex %>%
  ungroup() %>%
  mutate(age = case_when(age == "0 to 4" ~ "00 to 04",
                         age == "5 to 9" ~ "05 to 14",
                         age == "10 to 14" ~ "05 to 14",
                         age == "15 to 19" ~ "15 to 24",
                         age == "20 to 24" ~ "15 to 24",
                         age == "25 to 29" ~ "25 to 34",
                         age == "30 to 34" ~ "25 to 34",
                         age == "35 to 39" ~ "35 to 44",
                         age == "40 to 44" ~ "35 to 44",
                         age == "45 to 49" ~ "45 to 59",
                         age == "50 to 54" ~ "45 to 59",
                         age == "55 to 59" ~ "45 to 59",
                         TRUE ~ "60 & up")) %>%
  group_by(year, age, sex) %>%
  mutate(value = sum(value)) %>%
  ungroup()



m_age_sex <- lm(value ~ splines::ns(year, knots = 3)*age*sex, data = age_sex)

summary(m_age_sex)

age_levels <- age_sex %>% select(age) %>% distinct() %>% pull() 

age_sex_nd <- 
  crossing(
    age=age_levels,
    sex=c("male", "female"),
    year = 1950:1963
  )

pred_pops <- age_sex_nd %>% modelr::add_predictions(m_age_sex)

pred_pops %>%
  ggplot(aes(x=year, y=pred, colour=age)) +
  geom_line() +
  geom_point() +
  facet_grid(sex~.) +
  scale_y_continuous(labels = comma, limits = c(0, 125000))

#How well do they match up with our overall populations?
pred_pops %>%
  group_by(year) %>%
  summarise(sum_pred_pop = sum(pred)) %>%
  right_join(overall_pops) %>%
  select(year, sum_pred_pop, population_without_inst_ship, total_population) %>%
  pivot_longer(cols = c(sum_pred_pop, population_without_inst_ship, total_population)) %>%
  ggplot(aes(x=year, y=value, colour=name)) +
  geom_point() +
  scale_y_continuous(labels = comma, limits = c(800000, 1250000))

pred_pops %>%
  group_by(year, sex) %>%
  summarise(sum = sum(pred)) %>%
  group_by(year) %>%
  mutate(sex_ratio = first(sum)/last(sum))
```


What percentage of adults (15+ participated in the intervention in 1957)?

```{r}

pred_pops %>%
  ungroup() %>%
  filter(year==1957) %>%
  filter(age != "00 to 04",
         age != "05 to 14") %>%
  summarise(total_pop = sum(pred)) %>%
  mutate(cxr_screened = 714915) %>%
  mutate(pct_pop_cxr_screened = percent(cxr_screened/total_pop))

pred_pops %>%
  ungroup() %>%
  filter(year==1957) %>%
  filter(age != "00 to 04",
         age != "05 to 14") %>%
  summarise(total_pop = sum(pred), .by=sex) %>%
  mutate(cxr_screened = c(340474, 281875)) %>%
  mutate(pct_pop_cxr_screened = percent(cxr_screened/total_pop))


```



Population pyramids

```{r}

label_abs <- function(x) {
  comma(abs(x))
}


pred_pops %>%
  ungroup() %>%
  group_by(year) %>%
  mutate(year_pop = sum(pred),
         age_sex_pct = percent(pred/year_pop, accuracy=0.1)) %>%
  mutate(sex = case_when(sex=="male" ~ "Male",
                         sex=="female" ~ "Female")) %>%
  ggplot(
    aes(x = age, fill = sex, 
        y = ifelse(test = sex == "Female",yes = -pred, no = pred))) + 
  geom_bar(stat = "identity") +
  geom_text(aes(label = age_sex_pct),
            position= position_stack(vjust=0.5), colour="white", size=2.5) +
  facet_wrap(year~., ncol=7) +
  coord_flip() +
  scale_y_continuous(labels = label_abs) +
  scale_fill_manual(values = c("mediumseagreen", "purple"), name="") +
  theme_ggdist() +
  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust=0.5),
        legend.position = "bottom",
        panel.border = element_rect(colour = "grey78", fill=NA)) +
  labs(x="", y="") 


ggsave(here("figures/s3.png"), width=10)


```

Not perfect, but resonably good. But ahhhhh... the age groups don't align with the case notification age groups! Come back to think about this later.


### 4. Tuberculosis cases

Import the tuberculosis cases dataset


#### 4.1 Overall notifications

Overall, by year.

```{r}

cases_by_year <- read_xlsx("2023-11-28_glasgow-acf.xlsx", sheet = "by_year")

cases_by_year%>%
  mutate(across(where(is.numeric) & !(year),  ~comma(.))) %>%
  datatable()


#shift year to midpoint
cases_by_year <- cases_by_year %>%
  mutate(year2 = year+0.5)

```

Plot the overall number of case notified per year, by pulmonary and extra pulmonary classification.

```{r}

cases_by_year %>%
  select(-total_notifications, -year) %>%
  pivot_longer(cols = c(pulmonary_notifications, `non-pulmonary_notifications`)) %>%
  mutate(name = case_when(name == "pulmonary_notifications" ~ "Pulmonary TB",
                          name == "non-pulmonary_notifications" ~ "Extra-pulmonary TB")) %>%
  ggplot() +
  geom_area(aes(y=value, x=year2, group = name, fill=name), alpha=0.5) +
  geom_vline(aes(xintercept=acf_start), linetype=3) +
  geom_vline(aes(xintercept=acf_end), linetype=3) +
  scale_y_continuous(labels=comma) +
  scale_x_continuous(labels = year_labels,
                     breaks = year_labels) +
  scale_fill_brewer(palette = "Set1", name="") +
  labs(
    title = "Glasgow Corporation: Tuberculosis notifications",
    subtitle = "1950 to 1963, by TB disease classification",
    x = "Year",
    y = "Number of cases",
    caption = "Mass miniature X-ray campaign period between dashed lines (11th March-12th April 1957)"
  ) +
  theme_ggdist() +
  theme(legend.position = "bottom")
  

```

#### 4.2 Notifications by Division

Read in the datasets and merge together.

```{r}

#list all the sheets
all_sheets <- excel_sheets("2023-11-28_glasgow-acf.xlsx")

#get the ward sheets
ward_sheets <- enframe(all_sheets) %>%
  filter(grepl("by_ward", value)) %>%
  pull(value)


cases_by_ward_sex_year <- map_df(ward_sheets, ~read_xlsx(path = "2023-11-28_glasgow-acf.xlsx",
                                sheet = .))

cases_by_ward_sex_year %>%
  mutate(across(where(is.numeric) & !(year),  ~comma(.))) %>%
  datatable()

```

Aggregate together to get cases by division

```{r}

cases_by_division <- cases_by_ward_sex_year %>%
  left_join(ward_lookup) %>%
  group_by(division, year, tb_type) %>%
  summarise(cases = sum(cases, na.rm = TRUE))

#shift year to midpoint
cases_by_division <- cases_by_division %>%
  mutate(year2 = year+0.5) %>%
  ungroup()

cases_by_division  %>%
  select(-year2) %>%
  select(year, everything()) %>%
  mutate(across(where(is.numeric) & !(year),  ~comma(.))) %>%
  datatable()


cases_by_division %>%
  mutate(tb_type = case_when(tb_type == "Pulmonary" ~ "Pulmonary TB",
                          tb_type == "Non-Pulmonary" ~ "Extra-pulmonary TB")) %>%
  ggplot() +
  geom_area(aes(y=cases, x=year2, group = tb_type, fill=tb_type), alpha=0.5) +
  geom_vline(aes(xintercept=acf_start), linetype=3) +
  geom_vline(aes(xintercept=acf_end), linetype=3) +
  scale_y_continuous(labels=comma) +
  scale_x_continuous(labels = year_labels,
                     breaks = year_labels,
                     guide = guide_axis(angle = 90)) +
  facet_wrap(division~., scales = "free_y") +
  scale_fill_brewer(palette = "Set1", name="") +
  labs(
    title = "Glasgow Corporation: Tuberculosis notifications by Division",
    subtitle = "1950 to 1963, by TB disease classification",
    x = "Year",
    y = "Number of cases",
    caption = "Mass miniature X-ray campaign period between dashed lines (11th March-12th April 1957)\nNote: extra-pulmonary TB cases by Division/Ward not reported in 1962-1963"
  ) +
  theme_ggdist() +
  theme(legend.position = "bottom")

```

#### 4.3 Notifications by ward

```{r}


cases_by_ward <- cases_by_ward_sex_year %>%
  group_by(ward, year, tb_type) %>%
  summarise(cases = sum(cases, na.rm = TRUE)) %>%
  ungroup()

cases_by_ward %>%
  mutate(across(where(is.numeric) & !(year),  ~comma(.))) %>%
  select(year, everything()) %>%
  datatable()

#shift year to midpoint
cases_by_ward <- cases_by_ward %>%
  mutate(year2 = year+0.5)

cases_by_ward %>%
  mutate(tb_type = case_when(tb_type == "Pulmonary" ~ "Pulmonary TB",
                          tb_type == "Non-Pulmonary" ~ "Extra-pulmonary TB")) %>%
  ggplot() +
  geom_area(aes(y=cases, x=year2, group = tb_type, fill=tb_type), alpha=0.8) +
  geom_vline(aes(xintercept=acf_start), linetype=3) +
  geom_vline(aes(xintercept=acf_end), linetype=3) +
  scale_y_continuous(labels=comma) +
  scale_x_continuous(labels = year_labels,
                     breaks = year_labels,
                     guide = guide_axis(angle = 90)) +
  facet_wrap(ward~., scales = "free_y") +
  scale_fill_brewer(palette = "Set1", name="") +
  labs(
    title = "Glasgow Corporation: Tuberculosis notifications by Ward",
    subtitle = "1950 to 1963, by TB disease classification",
    x = "Year",
    y = "Number of cases",
    caption = "Mass miniature X-ray campaign period between dashed lines (11th March-12th April 1957)\nNote: extra-pulmonary TB cases by Division/Ward not reported in 1962-1963"
  ) +
  theme(legend.position = "bottom")


```

#### 4.4 Notifications by age and sex

As we don't have denominators, we will just model the change in counts.

```{r}

#list all the sheets
all_sheets <- excel_sheets("2023-11-28_glasgow-acf.xlsx")

#get the ward sheets
age_sex_sheets <- enframe(all_sheets) %>%
  filter(grepl("by_age_sex", value)) %>%
  pull(value)


cases_by_age_sex <- map_df(age_sex_sheets, ~read_xlsx(path = "2023-11-28_glasgow-acf.xlsx",
                                sheet = .))

cases_by_age_sex %>%
  mutate(across(where(is.numeric) & !(year),  ~comma(.))) %>%
  datatable()


```




### 5 TB incidence

#### 5.1 Overall TB incidence

Now calculate incidence per 100,000 population

Merge the notification and population denominator datasets together.

Here we need to include the whole population (with shipping and institutions) as they are included in the notifications.

```{r}

overall_inc <- overall_pops %>%
  left_join(cases_by_year)

overall_inc <- overall_inc %>%
  mutate(inc_pulm_100k = pulmonary_notifications/total_population*100000,
         inc_ep_100k = `non-pulmonary_notifications`/total_population*100000,
         inc_100k = total_notifications/total_population*100000)

overall_inc %>%
  select(year, inc_100k, inc_pulm_100k, inc_ep_100k) %>%
  mutate_at(.vars = vars(inc_100k, inc_pulm_100k, inc_ep_100k),
            .funs = funs(round)) %>%
  datatable()

```

```{r}

overall_inc %>%
  select(year2, inc_pulm_100k, inc_ep_100k) %>%
  pivot_longer(cols = c(inc_pulm_100k, `inc_ep_100k`)) %>%
  mutate(name = case_when(name == "inc_pulm_100k" ~ "Pulmonary TB",
                          name == "inc_ep_100k" ~ "Extra-pulmonary TB")) %>%
  ggplot() +
  geom_area(aes(y=value, x=year2, group = name, fill=name), alpha=0.5) +
  geom_vline(aes(xintercept=acf_start), linetype=3) +
  geom_vline(aes(xintercept=acf_end), linetype=3) +
  scale_y_continuous(labels=comma) +
  scale_x_continuous(labels = year_labels,
                     breaks = year_labels) +
  scale_fill_brewer(palette = "Set1", name="") +
  labs(
    title = "Glasgow Corporation: Tuberculosis case notification rate",
    subtitle = "1950 to 1963, by TB disease classification",
    x = "Year",
    y = "Case notification rate (per 100,000)",
    caption = "Mass miniature X-ray campaign period between dashed lines (11th March-12th April 1957)"
  ) +
  theme_ggdist() +
  theme(legend.position = "bottom")



```

#### 5.2 TB incidence by Division

```{r}

division_inc <- division_pops %>%
  left_join(cases_by_division)


division_inc <- division_inc %>%
  mutate(inc_100k = cases/total_population*100000)

division_inc %>%
  select(year, division, tb_type, inc_100k) %>%
  mutate_at(.vars = vars(inc_100k),
            .funs = funs(round)) %>%
  datatable()


```

```{r}

division_inc %>%
  mutate(tb_type = case_when(tb_type == "Pulmonary" ~ "Pulmonary TB",
                          tb_type == "Non-Pulmonary" ~ "Extra-pulmonary TB")) %>%
  ggplot() +
  geom_area(aes(y=inc_100k, x=year2, group = tb_type, fill=tb_type), alpha=0.5) +
  geom_vline(aes(xintercept=acf_start), linetype=3) +
  geom_vline(aes(xintercept=acf_end), linetype=3) +
  scale_y_continuous(labels=comma) +
  scale_x_continuous(labels = year_labels,
                     breaks = year_labels,
                     guide = guide_axis(angle = 90)) +
  facet_wrap(division~.) +
  scale_fill_brewer(palette = "Set1", name="") +
  labs(
    title = "Glasgow Corporation: Tuberculosis case notification rate, by Division",
    subtitle = "1950 to 1963, by TB disease classification",
    x = "Year",
    y = "Case notification rate (per 100,000)",
    caption = "Mass miniature X-ray campaign period between dashed lines (11th March-12th April 1957)\nNote: extra-pulmonary TB cases by Division/Ward not reported in 1962-1963"
  ) +
  theme_ggdist() +
  theme(legend.position = "bottom")



```

#### 5.2 TB incidence by Ward

Here we will filter out the institutions and harbour from the denominators, as we don't have reliable population denominators for them.

```{r}

ward_inc <- ward_pops %>%
  left_join(cases_by_ward)


ward_inc <- ward_inc %>%
  mutate(inc_100k = cases/population_without_inst_ship*100000)

ward_inc %>%
  select(year, ward, tb_type, inc_100k) %>%
  mutate_at(.vars = vars(inc_100k),
            .funs = funs(round)) %>%
  datatable()


```


```{r}

ward_inc %>%
  mutate(tb_type = case_when(tb_type == "Pulmonary" ~ "Pulmonary TB",
                          tb_type == "Non-Pulmonary" ~ "Extra-pulmonary TB")) %>%
  ggplot() +
  geom_area(aes(y=inc_100k, x=year2, group = tb_type, fill=tb_type), alpha=0.5) +
  geom_vline(aes(xintercept=acf_start), linetype=3) +
  geom_vline(aes(xintercept=acf_end), linetype=3) +
  scale_y_continuous(labels=comma) +
  scale_x_continuous(labels = year_labels,
                     breaks = year_labels,
                     guide = guide_axis(angle = 90)) +
  facet_wrap(ward~.) +
  scale_fill_brewer(palette = "Set1", name="") +
  labs(
    title = "Glasgow Corporation: Tuberculosis case notification rate, by Ward",
    subtitle = "1950 to 1963, by TB disease classification",
    x = "Year",
    y = "Incidence (per 100,000)",
    caption = "Mass miniature X-ray campaign period between dashed lines (11th March-12th April 1957)\nNote: extra-pulmonary TB cases by Division/Ward not reported in 1962-1963"
  ) +
  theme(legend.position = "bottom")




```

On a map

```{r, warning=FALSE}

st_as_sf(left_join(ward_inc, glasgow_wards_1951)) %>%
  filter(tb_type=="Pulmonary") %>%
  ggplot() +
  geom_sf(aes(fill=inc_100k)) +
  facet_wrap(year~., ncol = 7) +
  scale_fill_viridis_c(name="Case notification rate (per 100,000)",
                       option = "A") +
  theme_ggdist() +
  theme(legend.position = "top",
        legend.key.width = unit(2, "cm"),
        panel.border = element_rect(colour = "grey78", fill=NA)) +
  guides(fill=guide_colorbar(title.position = "top"))



```


### 6. TB Mortality

#### 6.1 Overall Mortality

Import the TB mortality data.

First, overall deaths. Note that in the original reports, we have a pulmonary TB death rate per million for all years, and numbers of pulmonary TB deaths for each year apart from 1950.

```{r}

#get the overall mortality sheets
deaths_sheets <- enframe(all_sheets) %>%
  filter(grepl("deaths", value)) %>%
  pull(value)


overall_deaths <- map_df(deaths_sheets, ~read_xlsx(path = "2023-11-28_glasgow-acf.xlsx",
                                sheet = .))

overall_deaths %>%
  mutate(across(where(is.numeric) & !(year),  ~comma(.))) %>%
  datatable()



```

Plot the raw numbers of pulmonary deaths

```{r}

overall_deaths %>%
  ggplot(aes(x=year, y=pulmonary_deaths)) +
  geom_line(colour = "#DE0D92") +
  geom_point(colour = "#DE0D92") +
  geom_vline(aes(xintercept=acf_start), linetype=3) +
  geom_vline(aes(xintercept=acf_end), linetype=3) +
  labs(y="Pulmonary TB deaths per year",
       x = "Year",
       title = "Numbers of pulmonary TB deaths",
       subtitle = "Glasgow, 1950-1963",
    caption = "Mass miniature X-ray campaign period between dashed lines (11th March-12th April 1957)\nNote: no data for 1950") +
  theme_ggdist() +
  theme(panel.border = element_rect(colour = "grey78", fill=NA))


```

Now the incidence of pulmonary TB death

```{r}
overall_deaths %>%
  ggplot(aes(x=year, y=pulmonary_death_rate_per_100k)) +
  geom_line(colour = "#4D6CFA") +
  geom_point(colour = "#4D6CFA") +
  geom_vline(aes(xintercept=acf_start), linetype=3) +
  geom_vline(aes(xintercept=acf_end), linetype=3) +
  scale_y_continuous(labels=comma) +
  scale_x_continuous(labels = year_labels,
                     breaks = year_labels) +
  labs(y="Annual incidence of death (per 100,000)",
       x = "Year",
    caption = "Mass miniature X-ray campaign period between dashed lines (11th March-12th April 1957)") +
  theme_ggdist() +
  theme(panel.border = element_rect(colour = "grey78", fill=NA))

ggsave(here("figures/s7.png"), width=10)

```


### 6. Table 1

Make Table 1 here, and save for publication.

```{r}

overall_pops %>% 
  select(year, total_population) %>%
  left_join(overall_inc %>%
              select(year, 
                     pulmonary_notifications, inc_pulm_100k,
                     `non-pulmonary_notifications`, inc_ep_100k,
                     total_notifications, inc_100k)) %>%
  left_join(overall_deaths %>%
              select(year,
                     pulmonary_deaths, pulmonary_death_rate_per_100k)) %>%
  mutate(across(where(is.numeric) & !(year),  ~round(., digits=1))) %>%
  mutate(across(where(is.numeric) & !(year),  ~comma(.))) 

```




### 7. Overall pulmonary TB model


#### 7.1 FIt the model and priors

First model will investigate the impact of mass miniature X-ray campaign on pulmonary TB case notification rate using an interrupted time series analysis.

Set up the data

```{r}

mdata1 <- overall_inc %>%
  mutate(acf_period = case_when(year %in% c(1950:1956) ~ "a. pre-acf",
                                year %in% c(1957) ~ "b. acf",
                                year %in% c(1958:1963) ~ "c. post-acf")) %>%
  mutate(y_num = 1:nrow(.)) %>%
  rename(extrapulmonary_notifications = `non-pulmonary_notifications`)

```


Work on the priors a bit

Basic prior

```{r}

basic_prior <- c(prior(normal(0, 1), class = Intercept),
                 prior(normal(0, 0.25), class = b))
```

Look at the mean and variance of counts (counts of pulmonary notifications are what we are predicting)

```{r}

#Mean of counts per year
mean(mdata1$pulmonary_notifications)
#variance of counts per year
var(mdata1$pulmonary_notifications)

```


Quite a bit of over-dispersion here, so negative binomial distribution might be a better choice of distributional family than Poisson.

Slightly more informative prior ("weakly informative" really)

```{r}

ggplot(data = tibble(x = seq(from = 500, to = 5000, by = 10)),
       aes(x = x, y = dgamma(x, shape = 2, rate = 0.001))) +
  geom_area(color = "transparent", 
            fill = "#DE0D92") +
  scale_x_continuous(NULL) +
  coord_cartesian(xlim = c(500, 5000)) +
  ggtitle(expression(brms~~gamma(2*", "*0.001)~shape~prior))


```

Fit a model with only priors

```{r}

winform_prior <- c(prior(normal(0, 1), class = Intercept),
                  prior(gamma(2, 0.001), class = shape),
                  prior(normal(0, 0.25), class = b))


m_pulmonary_prior <- brm(
  pulmonary_notifications ~ y_num + acf_period + acf_period:y_num +  offset(log(total_population)),
                  data = mdata1,
                  family = negbinomial(),
                  seed = 1234,
                  chains = 4, cores = 4, 
                  prior = winform_prior,
                  sample_prior = "only",
                  backend="cmdstanr",
                  save_pars = save_pars(all = TRUE),
                  warmup = 1000)

summary(m_pulmonary_prior)
conditional_effects(m_pulmonary_prior)

```

Now fit the model with the weakly informative priors


```{r}
m_pulmonary_overall <- brm(
  pulmonary_notifications ~ y_num + acf_period + acf_period:y_num +  offset(log(total_population)),
                  data = mdata1,
                  family = negbinomial(),
                  seed = 1234,
                  chains = 4, cores = 4, 
                  prior = winform_prior,
                  save_pars = save_pars(all = TRUE),
                  backend = "cmdstanr",
                  warmup = 1000)

summary(m_pulmonary_overall)
plot(m_pulmonary_overall)
pp_check(m_pulmonary_overall, type='ecdf_overlay')

```




#### 7.2 Summarise change in CNRs

Summarise the posterior

```{r}

f1b <- plot_counterfactual(model_data = mdata1, model = m_pulmonary_overall, population_denominator = total_population, outcome = inc_pulm_100k, grouping_var=NULL)
  
f1b
```

Make this into a figure

```{r}

f1a <- st_as_sf(left_join(ward_inc, glasgow_wards_1951)) %>%
  filter(tb_type=="Pulmonary") %>%
  ggplot() +
  geom_sf(aes(fill=inc_100k)) +
  facet_wrap(year~., ncol = 7) +
  scale_fill_viridis_c(name="Case notification rate (per 100,000)",
                       option = "A") +
  theme_ggdist() +
  theme(panel.border = element_rect(colour = "grey78", fill=NA),
        legend.position = "top",
        #legend.key.width = unit(2, "cm"),
        legend.title.align = 0.5) +
  guides(fill=guide_colorbar(title.position = "top"))

(f1a / f1b) + plot_annotation(tag_levels = "A")

ggsave(here("figures/f1.png"))

```


Summary of change in notifications

```{r}

summarise_change(model_data=mdata1, model=m_pulmonary_overall, population_denominator=total_population, grouping_var=NULL) %>%
  map(datatable)

```
(Alternative way - keep in for now)

```{r}

overall_immediate_draws <- mdata1 %>%
  select(year, year2, y_num, acf_period, total_population, pulmonary_notifications) %>%
  filter(year %in% c(1956,1957)) %>%
  add_epred_draws(m_pulmonary_overall) %>%
  mutate(inc_100k = .epred/total_population*100000) %>%
  group_by(.draw) %>%
  summarise(pct_change_immediate = (last(inc_100k) - first(inc_100k))/first(inc_100k)) %>%
  ungroup()

overall_post_draws <- mdata1 %>%
  select(year, year2, y_num, acf_period, total_population, pulmonary_notifications) %>%
  filter(year %in% c(1956,1958)) %>%
  add_epred_draws(m_pulmonary_overall) %>%
  mutate(inc_100k = .epred/total_population*100000) %>%
  group_by(.draw) %>%
  summarise(pct_change_post = (last(inc_100k) - first(inc_100k))/first(inc_100k)) %>%
  ungroup()


overall_slope_draws <- mdata1 %>%
  select(year, year2, y_num, acf_period, total_population, pulmonary_notifications) %>%
  filter(year %in% c(1950, 1956, 1958, 1963)) %>%
  add_epred_draws(m_pulmonary_overall) %>%
  mutate(inc_100k = .epred/total_population*100000) %>%
  ungroup() %>%
  mutate(n_years = length(year), .by=acf_period) %>%
  group_by(acf_period, .draw) %>%
  summarise(pct_change_slope = ((last(inc_100k) - first(inc_100k))/first(inc_100k))/n_years) %>%
  distinct() %>%
  pivot_wider(names_from = c(acf_period),
              values_from = pct_change_slope) %>%
  mutate(ratio_annual_slope = `c. post-acf` / `a. pre-acf`)


```

Correlation between immediate effect and post effect of ACF

```{r}

left_join(overall_immediate_draws, overall_post_draws) %>%
  ggplot(aes(x=pct_change_immediate, y=pct_change_post)) +
  geom_hdr(
    aes(fill = after_stat(probs)), 
    alpha = 1) +
  #geom_hdr_points(aes(colour = after_stat(probs)), size=0.5) +
  geom_smooth(method = "lm", se=FALSE, colour="black", linewidth=0.5) +
  stat_regline_equation(label.x = 0.25, label.y = -0.25, size=4) +
  scale_x_continuous(labels = percent,
                     breaks = pretty_breaks(n = 10)) +
  scale_y_continuous(labels = percent,
                     breaks = pretty_breaks(n = 5)) +
  scale_fill_viridis_d(option="E", name="") +
  labs(title="Correlation between immediate ACF impact and post-ACF case notification rate",
       y="Post intervention impact: ercentage change in CNR (1958 vs. 1956)",
       x="Immediate impact: percentage change in CNR (1957 vs. 1956)",
       caption="Boundaries are posterior desnity intervals from 4000 draws") +
  theme_ggdist() +
  theme(legend.position = "bottom",
        panel.border = element_rect(colour = "grey78", fill=NA))


```

Correlation between immediate effect and change in slope

```{r}

left_join(overall_immediate_draws, overall_slope_draws) %>%
  ggplot(aes(x=pct_change_immediate, y=ratio_annual_slope)) +
  geom_hdr(
    aes(fill = after_stat(probs)), 
    alpha = 1) +
  #geom_hdr_points(aes(colour = after_stat(probs)), size=0.5) +
  geom_smooth(method = "lm", se=FALSE, colour="black", linewidth=0.5) +
  #stat_regline_equation(label.x = 0.25, label.y = 0.02, size=4) +
  scale_x_continuous(labels = percent,
                     breaks = pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = pretty_breaks(n = 5),
                     limits = c(0, 10)) +
  scale_fill_viridis_d(option="E", name="") +
  labs(title="Correlation between immediate ACF impact and post-ACF case notification rate",
       y="Post intervention impact: Percentage change in CNR (1958 vs. 1956)",
       x="Immediate impact: percentage change in CNR (1957 vs. 1956)",
       caption="Points are draws from posteior distribution") +
  theme_ggdist() +
  theme(legend.position = "bottom",
        panel.border = element_rect(colour = "grey78", fill=NA))


```




#### 7.3 Compared to counterfactual

```{r}

overall_pulmonary_counterf <- calcuate_counterfactual(model_data = mdata1, model=m_pulmonary_overall, population_denominator = total_population)

overall_pulmonary_counterf$counter_post %>%
  mutate(across(c(cases_averted:cases_averted.upper, diff_inc100k:diff_inc100k.upper), number_format(accuracy = 0.1, big.mark = ","))) %>%
  mutate(across(c(rr_inc100k:rr_inc100k.upper), number_format(accuracy = 0.01))) %>%
  mutate(across(c(pct_change:pct_change.upper), percent, accuracy=0.1)) %>%
  datatable()


```

Total pulmonary TB cases averted between 1958 and 1963

```{r}

overall_pulmonary_counterf$counter_post_overall %>%
  mutate(across(c(cases_averted:cases_averted.upper), number_format(accuracy = 0.1, big.mark = ","))) %>%
  mutate(across(c(pct_change:pct_change.upper), percent, accuracy=0.1)) %>%
  datatable()


```



### 8. Extra-pulmonary TB notifications

#### 8.1 Fit the model

```{r, message=F, warning=FALSE}

m_extrap_overall <- brm(
  extrapulmonary_notifications ~ y_num + acf_period + acf_period:y_num + offset(log(total_population)),
                  data = mdata1,
                  family = poisson(),
                  seed = 1234,
                  chains = 4, cores = 4, 
                  prior = basic_prior,
                  save_pars = save_pars(all = TRUE),
                  backend = "cmdstanr",
                  warmup = 1000)

summary(m_extrap_overall)
plot(m_extrap_overall)
pp_check(m_extrap_overall, type='ecdf_overlay')

```


```{r}
plot_counterfactual(model_data = mdata1, model=m_extrap_overall, population_denominator = total_population, outcome=inc_ep_100k)
  
ggsave(here("figures/s6.png"), width=10)

```


#### 8.2 Summary of change

A. Percentage change in mortality, from 1956 to 1957 (i.e. immediate ACF effect)

```{r}

summarise_change(model_data=mdata1, model = m_extrap_overall, population_denominator = total_population) %>%
  map(datatable)

```



#### 8.3 Compared to counterfactual

```{r}

overall_ep_counterf <- calcuate_counterfactual(model_data = mdata1, model=m_extrap_overall, population_denominator = total_population)

overall_ep_counterf$counter_post %>%
  mutate(across(c(cases_averted:cases_averted.upper, diff_inc100k:diff_inc100k.upper), number_format(accuracy = 0.1, big.mark = ","))) %>%
  mutate(across(c(rr_inc100k:rr_inc100k.upper), number_format(accuracy = 0.01))) %>%
  mutate(across(c(pct_change:pct_change.upper), percent, accuracy=0.1)) %>%
  datatable()

```

Total extra pulmonary TB cases averted between 1958 and 1963

```{r}

overall_ep_counterf$counter_post_overall %>%
  mutate(across(c(cases_averted:cases_averted.upper), number_format(accuracy = 0.1, big.mark = ","))) %>%
  mutate(across(c(pct_change:pct_change.upper), percent, accuracy=0.1)) %>%
  datatable()


```


### 9. Ward level model

#### 9.1 Fit the model

```{r}

mdata2 <- ward_inc %>%
  filter(tb_type=="Pulmonary") %>%
  mutate(acf_period = case_when(year %in% c(1950:1956) ~ "a. pre-acf",
                                year %in% c(1957) ~ "b. acf",
                                year %in% c(1958:1963) ~ "c. post-acf")) %>%
  group_by(ward) %>%
  mutate(y_num = row_number()) %>%
  ungroup()



```

(Note the denominator without institutionalised people and "shipping"!)

```{r}
#weakly informative priors for multilevel model
basic_prior2 <- c(prior(normal(0, 1), class = Intercept),
                 prior(normal(0, 0.1), class = b),
                 prior(cauchy(0,5), class="sd"))


ggplot(data = tibble(x = seq(from = 0, to = 500, by = 10)),
       aes(x = x, y = dgamma(x, shape = 1, rate = 0.01))) +
  geom_area(color = "transparent", 
            fill = "#DE0D92") +
  scale_x_continuous(NULL) +
  scale_y_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = c(0, 500)) +
  ggtitle(expression(brms~~gamma(1*", "*0.01)~shape~prior))

winform_prior2 <- c(prior(normal(0, 1), class = Intercept),
                  prior(gamma(1, 0.01), class = shape),
                  prior(normal(0, 1), class = b),
                  prior(cauchy(0,5), class="sd"),
                  prior(lkj(2), class="cor"))
```


```{r}

m_pulmonary_ward_prior <- brm(
  cases ~ y_num + acf_period + acf_period:y_num + (1 + y_num + acf_period + acf_period:y_num | ward) + offset(log(population_without_inst_ship)),
                  data = mdata2,
                  family = negbinomial(),
                  seed = 1234,
                  chains = 4, cores = 4, 
                  prior = winform_prior2,
                  save_pars = save_pars(all = TRUE),
                  sample_prior = "only",
                  backend = "cmdstanr",
                  warmup = 1000)

conditional_effects(m_pulmonary_ward_prior)


```

Now fit the model with data

```{r}
m_pulmonary_ward <- brm(
  cases ~ y_num + acf_period + acf_period:y_num + (1 + y_num + acf_period + acf_period:y_num | ward) + offset(log(population_without_inst_ship)),
                  data = mdata2,
                  family = negbinomial(),
                  seed = 1234,
                  chains = 4, cores = 4, 
                  prior = winform_prior2,
                  backend = "cmdstanr")

summary(m_pulmonary_ward)
plot(m_pulmonary_ward)
pp_check(m_pulmonary_ward, type='ecdf_overlay')


```


```{r}

plot_counterfactual(model_data = mdata2, model=m_pulmonary_ward, outcome = inc_100k, population_denominator = population_without_inst_ship, grouping_var = ward, ward)
  
ggsave(here("figures/s4.png"), width=10, height=12)

```

#### 9.2 Summary of change

A. percentage increase in CNR, from 1956 to 1957 (i.e. immediate ACF effect)

```{r}

ward_change <- summarise_change(model_data = mdata2, model = m_pulmonary_ward, population_denominator = population_without_inst_ship, grouping_var=ward) 

ward_change %>%
  map(datatable)

```

As a supplementary figure

```{r}
  
ward_change$immediate_change %>%
  arrange(acf_inc100k_rr) %>%
  ggplot() +
  geom_pointrange(aes(y=acf_inc100k_rr, ymin=acf_inc100k_rr.lower, ymax=acf_inc100k_rr.upper, 
                      x=fct_reorder(ward, acf_inc100k_rr),
                      colour = acf_inc100k_rr)) +
  geom_hline(aes(yintercept=1), linetype=2) +
  coord_flip() +
  scale_colour_viridis_b(option = "D") +
  scale_y_continuous(limits = c(0.8,3.0)) +
  labs(x="",
       y="Relative posterior predicted case notification rate (per 100,000; 95% UI)\nACF (1957) vs. Before ACF (1956)") +
  theme_ggdist() +
  theme(legend.position = "none",
        panel.border = element_rect(colour = "grey78", fill=NA))
  
ggsave(here("figures/s5.png"))

```


```{r}

ward_change$post_change %>%
  arrange(acf_inc100k_rr) %>%
  ggplot() +
  geom_pointrange(aes(y=acf_inc100k_rr, ymin=acf_inc100k_rr.lower, ymax=acf_inc100k_rr.upper, 
                      x=fct_reorder(ward, acf_inc100k_rr),
                      colour = acf_inc100k_rr)) +
  geom_hline(aes(yintercept=1), linetype=2) +
  coord_flip() +
  scale_colour_viridis_b(option = "D") +
  #scale_y_continuous(limits = c(0.8,3.0)) +
  labs(x="",
       y="Relative posterior predicted case notification rate (per 100,000; 95% UI)\nAfter ACF (1958) vs. Before ACF (1956)") +
  theme_ggdist() +
  theme(legend.position = "none",
        panel.border = element_rect(colour = "grey78", fill=NA))

ggsave(here("figures/s6.png"))


```


percentage change = (final value - initial value) / initial value

```{r}

ward_change$slope_change %>%
  ggplot() +
  geom_pointrange(aes(y=pct_change_epred_overall , ymin=pct_change_lower_overall , ymax=pct_change_upper_overall ,
                      group=acf_period, colour=acf_period,
                      x = fct_reorder(ward, pct_change_epred_overall ))) +
  coord_flip() +
  scale_y_continuous(labels =percent) +
  scale_colour_manual(values = c("#DE0D92", "#4D6CFA")) +
  #scale_y_continuous(limits = c(0.8,3.0)) +
  labs(title="Intervention effect of mass miniature chest X-ray in Glasgow",
       subtitle="By municipal ward",
       x="",
       y="Mean annual rate of change in case notification rate (95% CrI)\n Before ACF (1950-1956) vs. after ACF (1958-1963)") +
  theme_ggdist() +
  theme(panel.border = element_rect(colour = "grey78", fill=NA))

```

Suggestion from Pete D 2024-02-02: Try plotting these on choropleth maps


```{r}

bind_rows(
  (glasgow_wards_1951 %>%
     left_join(ward_change$immediate_change) %>%
     mutate(estimate = "Immediate effect")),
  (glasgow_wards_1951 %>%
     left_join(ward_change$post_change) %>%
     mutate(estimate = "Post ACF effect")),
  (glasgow_wards_1951 %>%
     left_join(ward_change$post_change) %>%
     mutate(estimate = "Slope effect"))
  ) %>%
  select(ward, acf_inc100k_rr, estimate) %>%
  mutate(ward = case_when(ward=="Shettleston and Tollcross" ~ "Shettleston and\nTollcross",
                          ward=="Partick (West)" ~ "Partick\n(West)",
                          ward=="Partick (East)" ~ "Partick\n(East)",
                          ward=="North Kelvin" ~ "North\nKelvin",
                          ward=="Kinning Park" ~ "Kinning\nPark",
                          TRUE ~ ward)) %>%
  split(.$estimate) %>%
  map(~ ggplot(.) +
  geom_sf(aes(fill=acf_inc100k_rr)) +
  geom_sf_label(aes(label = ward), size=1.5, fill=NA, label.size = NA, colour="black", family = "Segoe UI") +
  scale_fill_viridis_c() +
  theme_ggdist() +
  theme(panel.border = element_rect(colour = "grey78", fill=NA)) +
  labs(y="", x="",
       fill="RR")) %>%
    cowplot::plot_grid(plotlist = ., ncol = 3,
                       labels = c('A: Immediate effect (1957 vs. 1956)', 
                                  'B: Post-ACF effect (1958 vs. 1956)', 
                                  'C: Slope change effect (1958-63 vs. 1950-56'))

ggsave(here("figures/ward_effects.png"), width = 16)
  

```




(Alternative figure - keep in for the minute)

Is there any correlation between immediate increase and a) post-intervention (1958) effect, and b) post intervention slope (1958-1963)

Try a different way with the full distribution of posteriors


```{r}

#relative rate (1957 vs. 1956)
ward_immediate_draws <- mdata2 %>%
  select(year, year2, y_num, acf_period, population_without_inst_ship, cases, ward) %>%
  filter(year %in% c(1956,1957)) %>%
  add_epred_draws(m_pulmonary_ward) %>%
  mutate(inc_100k = .epred/population_without_inst_ship*100000) %>%
  group_by(.draw, ward) %>%
  summarise(rr_1957_vs_1956 = (last(inc_100k))/first(inc_100k)) %>%
  arrange(ward) %>%
  ungroup()

ward_immediate_draws %>%
  ggplot() +
  geom_histogram(aes(x=rr_1957_vs_1956, fill=ward, colour=ward), alpha=0.75) +
  facet_wrap(ward~.) +
  theme_ggdist() +
  theme(legend.position = "none")

#relative rate (1957 vs. mean(1950-1965))
ward_immediate_draws_expanded <- mdata2 %>%
  select(year, year2, y_num, acf_period, population_without_inst_ship, cases, ward) %>%
  filter(year %in% c(1950:1957)) %>%
  add_epred_draws(m_pulmonary_ward) %>%
  group_by(.draw, ward, acf_period) %>%
  summarise(.epred = mean(.epred),
            population_without_inst_ship = population_without_inst_ship) %>%
  group_by(.draw, ward) %>%
  mutate(inc_100k = .epred/population_without_inst_ship*100000) %>%
  summarise(rr_1957_vs_mean_1950_1956 = (last(inc_100k))/first(inc_100k)) %>%
  arrange(ward) %>%
  ungroup()

ward_immediate_draws_expanded %>%
  ggplot() +
  geom_histogram(aes(x=rr_1957_vs_mean_1950_1956, fill=ward, colour=ward), alpha=0.75) +
  facet_wrap(ward~.) +
  theme_ggdist() +
  theme(legend.position = "none")

#relative rate (1958 vs. 1956)
ward_post_draws <- mdata2 %>%
  select(year, year2, y_num, acf_period, population_without_inst_ship, cases, ward) %>%
  filter(year %in% c(1956,1958)) %>%
  add_epred_draws(m_pulmonary_ward) %>%
  mutate(inc_100k = .epred/population_without_inst_ship*100000) %>%
  group_by(.draw, ward) %>%
  summarise(rr_1958_vs_1956 = (last(inc_100k))/first(inc_100k)) %>%
  arrange(ward) %>%
  ungroup()

ward_post_draws %>%
  ggplot() +
  geom_histogram(aes(x=rr_1958_vs_1956, fill=ward, colour=ward), alpha=0.75) +
  facet_wrap(ward~.) +
  theme_ggdist() +
  theme(legend.position = "none")


#relative rate (1958 vs. mean(1950-1956))
ward_post_draws_expanded <- mdata2 %>%
  select(year, year2, y_num, acf_period, population_without_inst_ship, cases, ward) %>%
  filter(year %in% c(1950:1956, 1958)) %>%
  add_epred_draws(m_pulmonary_ward) %>%
  group_by(.draw, ward, acf_period) %>%
  summarise(.epred = mean(.epred),
            population_without_inst_ship = population_without_inst_ship) %>%
  group_by(.draw, ward) %>%
  mutate(inc_100k = .epred/population_without_inst_ship*100000) %>%
  summarise(rr_1958_vs_mean_1950_1956 = (last(inc_100k))/first(inc_100k)) %>%
  arrange(ward) %>%
  ungroup()

ward_post_draws_expanded %>%
  ggplot() +
  geom_histogram(aes(x=rr_1958_vs_mean_1950_1956, fill=ward, colour=ward), alpha=0.75) +
  facet_wrap(ward~.) +
  theme_ggdist() +
  theme(legend.position = "none")


#Was a greater relative increase in CNR associated with greater reductions post ACF
left_join(ward_immediate_draws, ward_post_draws) %>%
  ggplot(aes(y=rr_1958_vs_1956, x=rr_1957_vs_1956)) +
  geom_hdr_points(size=0.1) +
  geom_smooth(method = "lm", se=FALSE, colour="black", linewidth=0.5) +
  facet_wrap(ward~.)

left_join(ward_immediate_draws_expanded, ward_post_draws) %>%
  ggplot(aes(y=rr_1958_vs_1956, x=rr_1957_vs_mean_1950_1956)) +
  geom_hdr_points(size=0.1) +
  geom_smooth(method = "lm", se=FALSE, colour="black", linewidth=0.5) +
  facet_wrap(ward~.)

left_join(ward_immediate_draws, ward_post_draws_expanded) %>%
  ggplot(aes(y=rr_1958_vs_mean_1950_1956, x=rr_1957_vs_1956)) +
  geom_hdr_points(size=0.1) +
  geom_smooth(method = "lm", se=FALSE, colour="black", linewidth=0.5) +
  facet_wrap(ward~.)

left_join(ward_immediate_draws_expanded, ward_post_draws_expanded) %>%
  ggplot(aes(y=rr_1958_vs_mean_1950_1956, x=rr_1957_vs_mean_1950_1956)) +
  geom_hdr_points(size=0.1) +
  geom_smooth(method = "lm", se=FALSE, colour="black", linewidth=0.5) +
  facet_wrap(ward~.)

#regardless of comparison period, it looks like no, and in fact correlation is in opposite direction
#does this make sense?
#I guess that there is probably too short of a period to have an immediate effect on transmission
#Therefore, were cases just "brought forward in time" with ACF
#i.e. diagnosed in 1957 instead of 1958
#but the wards with the greatest increase in CNRs in 1957 probably had the *worst* TB transmission
#Therefore, what we are capturing here is probably:
#high transmission wards -> lots of cases detected -> but still have a higher CNR post intervention compared to lower transmission wards?
#TBD
#And more interesting might be the effect on slopes. i.e. did we alter the trajectory?
```


```{r}


ward_slope_draws <- mdata2 %>%
  select(year, year2, y_num, acf_period, population_without_inst_ship, cases, ward) %>%
  filter(year %in% c(1950, 1956, 1958, 1963)) %>%
  add_epred_draws(m_pulmonary_ward) %>%
  mutate(inc_100k = .epred/population_without_inst_ship*100000) %>%
  ungroup() %>%
  mutate(n_years = case_when(acf_period=="a. pre-acf" ~ 7,
                             acf_period=="c. post-acf" ~ 6)) %>%
  group_by(.draw, ward, acf_period) %>%
  summarise(slope = (last(inc_100k)/first(inc_100k)))

ward_slope_draws %>%
  ggplot() +
  geom_density(aes(x=slope, fill=acf_period),alpha=0.5) +
  facet_wrap(ward~.)

ward_slope_draws_rr <- ward_slope_draws %>%
  group_by(ward, .draw) %>%
  summarise(rr_slope = last(slope)/first(slope))

#correlation between slope rate ratio (post vs pre) and magnitude of ACF effect
#a) with magnitude compared to 1956 only
left_join(ward_immediate_draws, ward_slope_draws_rr) %>%
  ggplot(aes(y=rr_slope, x=rr_1957_vs_1956)) +
  geom_hdr_points(size=0.1) +
  geom_smooth(method = "lm", se=FALSE, colour="black", linewidth=0.5) +
  facet_wrap(ward~.)

#b) with magnitude compared to mean of 1950-1956 only
left_join(ward_immediate_draws_expanded, ward_slope_draws_rr) %>%
  ggplot(aes(y=rr_slope, x=rr_1957_vs_mean_1950_1956)) +
  geom_hdr_points(size=0.1) +
  geom_smooth(method = "lm", se=FALSE, colour="black", linewidth=0.5) +
  facet_wrap(ward~.)


```


Correlation between immediate effect and post effect of ACF

```{r}

# 
# left_join(ward_immediate_draws, ward_post_draws) %>%
#   ggplot(aes(x=pct_change_immediate, y=pct_change_post, group=ward)) +
#   geom_hdr_points(size=0.1) +
#   geom_smooth(method = "lm", se=FALSE, colour="black", linewidth=0.5) +
#   stat_regline_equation(label.x = 0, label.y = 0.25, size=4) +
#   scale_colour_scico_d(palette = "lipari", name = "Posterior probability") +
#   scale_x_continuous(labels = percent) +
#   scale_y_continuous(labels = percent) +
#   labs(title="Correlation between immediate ACF impact and post-ACF case notification rate",
#        y="Post intervention impact: ercentage change in CNR (1958 vs. 1956)",
#        x="Immediate impact: percentage change in CNR (1957 vs. 1956)",
#        caption="Points are draws from posteior distribution") +
#   theme_ggdist() +
#   theme(legend.position = "bottom",
#         panel.border = element_rect(colour = "grey78", fill=NA)) +
#   facet_wrap(ward~.)


```

Correlation between immediate effect and change in slope

```{r}

# left_join(ward_immediate_draws, ward_slope_draws) %>%
#   ggplot(aes(x=pct_change_immediate, y=ratio_annual_slope, group=ward)) +
#   geom_hdr_points(size=0.1) +
#   geom_smooth(method = "lm", se=FALSE, colour="black", linewidth=0.5) +
#   #stat_regline_equation(label.x = 0, label.y = 0.02, size=4) +
#   scale_colour_scico_d(palette = "lipari", name = "Posterior probability") +
#   scale_x_continuous(labels = percent) +
#   scale_y_continuous(limits = c(0, 10)) +
#   labs(title="Correlation between immediate ACF impact and post-ACF case notification rate",
#        y="Post intervention impact: Percentage change in CNR (1958 vs. 1956)",
#        x="Immediate impact: percentage change in CNR (1957 vs. 1956)",
#        caption="Points are draws from posteior distribution") +
#   theme_ggdist() +
#   theme(legend.position = "bottom",
#         panel.border = element_rect(colour = "grey78", fill=NA)) +
#   facet_wrap(ward~.)


```


Join these together with the overall estimates to make a single figure for showing impact

```{r}

# f2_data <- 
#   left_join(overall_immediate_draws, overall_post_draws) %>%
#   left_join(overall_slope_draws) %>%
#   mutate(level = "overall",
#          ward = "Glasgow") %>%
#   bind_rows(
#     left_join(ward_immediate_draws, ward_post_draws) %>%
#   left_join(ward_slope_draws) %>%
#   mutate(level = "ward")
#   )
# 
# f2a <- f2_data %>% 
#   ggplot(aes(x=pct_change_immediate, y=pct_change_post, group=ward)) +
#   geom_hdr_points(size=0.1) +
#   geom_smooth(method = "lm", se=FALSE, colour="black", linewidth=0.5) +
#   stat_regline_equation(label.x = 0, label.y = 0.12, size=4) +
#   scale_colour_scico_d(palette = "lipari", name = "Posterior probability density") +
#   scale_x_continuous(labels = percent) +
#   scale_y_continuous(labels = percent) +
#   labs(y="Post-intervention impact: Percentage change in CNR (1958 vs. 1956)",
#        x="Immediate impact: percentage change in CNR (1957 vs. 1956)") +
#   theme_ggdist() +
#   theme(legend.position = "bottom",
#         panel.border = element_rect(colour = "grey78", fill=NA)) +
#   facet_wrap(fct_relevel(ward,
#                          "Glasgow",
#                          after=0)~., ncol = 5) + 
#   guides(colour = guide_legend(override.aes = list(size=4)))
# 
# f2a
# 
# f2b <- f2_data %>% 
#   ggplot(aes(x=pct_change_immediate, y=ratio_annual_slope, group=ward)) +
#   geom_hdr_points(size=0.1) +
#   geom_smooth(method = "lm", se=FALSE, colour="black", linewidth=0.5) +
#   stat_regline_equation(label.x = 1.2, label.y = 12, size=4) +
#   scale_x_continuous(labels = percent,
#                      breaks = pretty_breaks(n = 4)) +
#   scale_y_continuous(breaks = pretty_breaks(n = 4),
#                 limits = c(0,15)) +
#   scale_fill_viridis_d(option="E") +
#   labs(y="Post-intervention impact: Relative change in annual CNR slope (1958-1963 vs. 1950-1956)",
#        x="Immediate impact: percentage change in CNR (1957 vs. 1956)",
#        colour="Posterior probability density") +
#   theme_ggdist() +
#   theme(legend.position = "bottom",
#         panel.border = element_rect(colour = "grey78", fill=NA)) +
#   facet_wrap(fct_relevel(ward,
#                          "Glasgow",
#                          after=0)~., ncol = 5) + 
#   guides(colour = guide_legend(override.aes = list(size=4)))
# 
# f2b
# 
# (f2a / f2b) + plot_annotation(tag_levels = 'A')

ggsave(here("figures/f2.png"), height=18, width=10)

```

#### 9.3 Compared to counterfactual

```{r}

ward_counterf <- calcuate_counterfactual(model_data = mdata2, model=m_pulmonary_ward, population_denominator = population_without_inst_ship, grouping_var=ward)

ward_counterf %>%
  map(datatable)


```

Total pulmonary TB cases averted between 1958 and 1963

```{r}

ward_averted <- ward_counterf$counter_post %>%
  summarise(across(c(cases_averted, cases_averted.lower, cases_averted.upper), sum), .by=ward) %>%
  mutate_if(is.double, ~ scales::number(x = ., accuracy = 0.1, big.mark = ",")) %>%
  mutate(cases_averted_txt = glue::glue("{cases_averted}\n({cases_averted.lower}-{cases_averted.upper})")) %>%
  select(ward, cases_averted_txt)

ward_averted %>% datatable()

```

Add the numbers averted for each ward to the figure

```{r}

plot_counterfactual(model_data = mdata2, model=m_pulmonary_ward, outcome = inc_100k, population_denominator = population_without_inst_ship, grouping_var = ward, ward) +
  geom_text(data=ward_averted, aes(x=1961, y=500, label=cases_averted_txt), size=3)
  
ggsave(here("figures/s4.png"), width=14, height=12)



```




### 10. Age-sex model

#### 10.1 FIt the model

Fit the model

(Not rewritten the functions for this yet)

```{r}

mdata3 <- cases_by_age_sex %>%
  filter(tb_type=="Pulmonary") %>%
  mutate(acf_period = case_when(year %in% c(1950:1956) ~ "a. pre-acf",
                                year %in% c(1957) ~ "b. acf",
                                year %in% c(1958:1963) ~ "c. post-acf")) %>%
  mutate(year2 = year+0.5) %>%
  group_by(age, sex) %>%
  mutate(y_num = row_number()) %>%
  ungroup()

winform_prior3 <- c(prior(normal(0, 1), class = Intercept),
                  #prior(gamma(0.5, 0.01), class = shape),
                  prior(normal(0, 1), class = b),
                  prior(cauchy(0,5), class="sd"),
                  prior(lkj(2), class="cor"))


m_age_sex <- brm(
  cases ~ y_num + (acf_period)*(age*sex) + (acf_period:y_num)*(age*sex),
                  data = mdata3,
                  family = negbinomial(),
                  seed = 1234,
                  chains = 4, cores = 4, 
                  prior = basic_prior,
                  backend = "cmdstanr")

summary(m_age_sex)
plot(m_age_sex)
pp_check(m_age_sex, type='ecdf_overlay')


```

Summarise posterior

```{r}

#posterior draws, and summarise
age_sex_summary <- mdata3 %>%
  select(year, year2, y_num, acf_period, age, sex) %>%
  add_epred_draws(m_age_sex) %>%
  group_by(year2, acf_period, age, sex) %>%
  mean_qi() %>%
  mutate(acf_period = case_when(acf_period=="a. pre-acf" ~ "Before Intervention",
                                acf_period=="c. post-acf" ~ "Post Intervention"))

#create the counterfactual (no intervention), and summarise
age_sex_counterfact <- 
  tibble(year = mdata3$year,
         year2 = mdata3$year2,
         y_num = mdata3$y_num,
         age = mdata3$age,
         sex = mdata3$sex,
         acf_period = factor("a. pre-acf")) %>%
  add_epred_draws(m_age_sex) %>%
  group_by(year2, acf_period, age, sex) %>%
  mean_qi() %>%
  mutate(acf_period = case_when(acf_period=="a. pre-acf" ~ "Before Intervention",
                                acf_period=="c. post-acf" ~ "Post Intervention")) %>%
  ungroup() %>%
  mutate(age = case_when(age=="00_05" ~ "0 to 5y",
                         age=="06_15" ~ "06 to 15y",
                         age=="16_25" ~ "16 to 25y",
                         age=="26_35" ~ "26 to 35y",
                         age=="36_45" ~ "36 to 45y",
                         age=="46_55" ~ "46 to 55y",
                         age=="56_65" ~ "56 to 65y",
                         age=="65+" ~ "65 & up y")) %>%
  mutate(sex = case_when(sex== "M" ~ "Male",
                         sex== "F" ~ "Female")) 



age_sex_summary %>%
  ungroup() %>%
  mutate(age = case_when(age=="00_05" ~ "0 to 5y",
                         age=="06_15" ~ "06 to 15y",
                         age=="16_25" ~ "16 to 25y",
                         age=="26_35" ~ "26 to 35y",
                         age=="36_45" ~ "36 to 45y",
                         age=="46_55" ~ "46 to 55y",
                         age=="56_65" ~ "56 to 65y",
                         age=="65+" ~ "65 & up y")) %>%
  mutate(sex = case_when(sex== "M" ~ "Male",
                         sex== "F" ~ "Female")) %>%
  ggplot() +
  geom_ribbon(aes(ymin=.epred.lower, ymax=.epred.upper, x=year2, group = acf_period, fill=acf_period), alpha=0.5) +
  geom_ribbon(data = age_sex_counterfact %>% filter(year>=1956), 
              aes(ymin=.epred.lower, ymax=.epred.upper, x=year2, fill="Counterfactual"), alpha=0.5) +
  geom_line(data = age_sex_counterfact %>% filter(year>=1956), 
              aes(y=.epred, x=year2, colour="Counterfactual")) +
  geom_line(aes(y=.epred, x=year2, group=acf_period,  colour=acf_period)) +
  geom_point(data = mdata3 %>%
  ungroup() %>%
  mutate(age = case_when(age=="00_05" ~ "0 to 5y",
                         age=="06_15" ~ "06 to 15y",
                         age=="16_25" ~ "16 to 25y",
                         age=="26_35" ~ "26 to 35y",
                         age=="36_45" ~ "36 to 45y",
                         age=="46_55" ~ "46 to 55y",
                         age=="56_65" ~ "56 to 65y",
                         age=="65+" ~ "65 & up y")) %>%
  mutate(sex = case_when(sex== "M" ~ "Male",
                         sex== "F" ~ "Female")) , aes(y=cases, x=year2, shape=acf_period), size=2) +
  geom_vline(aes(xintercept=acf_end), linetype=3) +
  ggh4x::facet_grid2(age~sex, scales = "free_y", independent = "y") +
  theme_ggdist() +
  scale_y_continuous(labels=comma) +
  scale_x_continuous(labels = year_labels,
                     breaks = year_labels,
                     guide = guide_axis(angle = 90)) +
  scale_fill_manual(values = c("#DE0D92", "grey50", "#4D6CFA") , name="") +
  scale_colour_manual(values = c("#DE0D92", "grey50", "#4D6CFA") , name="") +
  scale_shape_discrete(name="") +
  labs(
    x = "Year",
    y = "Case notifications (n)",
    caption = "Mass miniature X-ray campaign period between dashed lines (11th March-12th April 1957)"
  ) +
  theme(legend.position = "bottom",
        panel.border = element_rect(colour = "grey78", fill=NA),
        title = element_text(size=14),
        axis.text = element_text(size=14),
        legend.text = element_text(size=12)) +
  guides(shape="none")
  
ggsave(here("figures/s7.png"), height=10)

```

#### 10.2 Summary of impact of intervention

1. percentage increase in CNR, from 1956 to 1957 (i.e. immediate ACF effect)

```{r}

nd <- mdata3 %>%
  filter(year %in% c(1956:1957)) %>%
  select(acf_period, y_num, age, sex)


age_sex_impact_out <- 
  add_epred_draws(m_age_sex,
                newdata=nd) %>%
  ungroup() %>%
  select(acf_period, .epred, age, sex) %>%
  pivot_wider(names_from = acf_period,
              values_from = .epred,
              values_fn = list) %>%
  unnest() %>%
  rename(pre_epred = 3,
         post_epred = 4) %>%
  mutate(acf_diff = post_epred-pre_epred,
         acf_rr = post_epred/pre_epred) %>%
  group_by(age, sex) %>%
  mean_qi(acf_diff, acf_rr) 

age_sex_impact_out %>%
  mutate_if(is.double, ~ scales::number(x = ., accuracy = 0.01, big.mark = ",")) %>%
  datatable()
  
f3a <- age_sex_impact_out %>%
  mutate(sex = case_when(sex=="M" ~ "Male",
                         sex=="F" ~ "Female")) %>%
  mutate(age = case_when(age=="00_05" ~ "0 to 5y",
                         age=="06_15" ~ "06 to 15y",
                         age=="16_25" ~ "16 to 25y",
                         age=="26_35" ~ "26 to 35y",
                         age=="36_45" ~ "36 to 45y",
                         age=="46_55" ~ "46 to 55y",
                         age=="56_65" ~ "56 to 65y",
                         age=="65+" ~ "65 & up y")) %>%
  ggplot() +
  geom_pointrange(aes(y=acf_rr, ymin=acf_rr.lower, ymax=acf_rr.upper, group=sex, 
                      x=age,
                      colour = sex),
                  position = position_dodge(width = 0.25)) +
  geom_hline(aes(yintercept=1), linetype=2) +
  scale_colour_manual(values = c("purple", "darkorange"), name="") +
  labs(x="",
       y="Relative notifications (95% UI)\nACF (1957) vs. Before ACF (1956)") +
  theme_ggdist() +
  theme(legend.position = "bottom",
        panel.border = element_rect(colour = "grey78", fill=NA))
  
  

```


2. Change from pre-ACF period (1956), to first year post-ACF (1958)


```{r}

nd <- mdata3 %>%
  filter(year %in% c(1956,1958)) %>%
  select(acf_period, y_num, age, sex)

#Do it with calculating incidence, then sumamrising.
age_sex_impact2 <-add_epred_draws(m_age_sex,
                newdata=nd) %>%
  ungroup() %>%
  select(acf_period, .epred, age, sex) %>%
  pivot_wider(names_from = acf_period,
              values_from = .epred,
              values_fn = list) %>%
  unnest() %>%
  rename(pre_epred = 3,
        post_epred = 4) %>%
  mutate(acf_diff = post_epred-pre_epred,
         acf_rr = post_epred/pre_epred) %>%
  group_by(age, sex) %>%
  mean_qi(acf_diff, acf_rr) 

age_sex_impact2 %>%
  mutate_if(is.double, ~ scales::number(x = ., accuracy = 0.01, big.mark = ",")) %>%
  datatable()

f3b <- age_sex_impact2 %>%  
  mutate(sex = case_when(sex=="M" ~ "Male",
                         sex=="F" ~ "Female")) %>%
  mutate(age = case_when(age=="00_05" ~ "0 to 5y",
                         age=="06_15" ~ "06 to 15y",
                         age=="16_25" ~ "16 to 25y",
                         age=="26_35" ~ "26 to 35y",
                         age=="36_45" ~ "36 to 45y",
                         age=="46_55" ~ "46 to 55y",
                         age=="56_65" ~ "56 to 65y",
                         age=="65+" ~ "65 & up y")) %>%
  ggplot() +
  geom_pointrange(aes(y=acf_rr, ymin=acf_rr.lower, ymax=acf_rr.upper, group=sex, 
                      x=age,
                      colour = sex),
                  position = position_dodge(width = 0.25)) +
  geom_hline(aes(yintercept=1), linetype=2) +
  scale_colour_manual(values = c("purple", "darkorange"), name="") +
  labs(x="",
       y="Relative notifications (95% UI)\nACF (1958) vs. Before ACF (1956)") +
  theme_ggdist() +
  theme(legend.position = "bottom",
        panel.border = element_rect(colour = "grey78", fill=NA))


```

3. Change in slope (i.e. difference in mean annual case notification rate pre-Intervention vs. post-intervention, by ward)

```{r}

age_sex_impact3 <- mdata3 %>%
  select(year, year2, y_num, acf_period, cases, age, sex) %>%
  filter(year!=1957) %>%
  add_epred_draws(m_age_sex) %>%
  group_by(year, age, sex, acf_period) %>%
  mean_qi(.epred) %>%
  ungroup() %>%
  mutate(n_years = length(year), .by=acf_period) %>%
  summarise(pct_change_epred_overall = (((last(.epred) - first(.epred))/first(.epred))),
            pct_change_lower_overall = (((last(.lower) - first(.lower))/first(.lower))),
            pct_change_upper_overall = (((last(.upper) - first(.upper))/first(.upper))),
    
            pct_change_epred_annual = (((last(.epred) - first(.epred))/first(.epred))/n_years),
            pct_change_lower_annual = (((last(.lower) - first(.lower))/first(.lower))/n_years),
            pct_change_upper_annual = (((last(.upper) - first(.upper))/first(.upper))/n_years),
            .by = c(acf_period, age, sex)) %>%
  distinct()


age_sex_impact3 %>%
  mutate_if(is.double, percent) %>%
  datatable()

f3c <- age_sex_impact3 %>%
  mutate(sex = case_when(sex=="M" ~ "Male",
                         sex=="F" ~ "Female")) %>%
  mutate(age = case_when(age=="00_05" ~ "0 to 5y",
                         age=="06_15" ~ "06 to 15y",
                         age=="16_25" ~ "16 to 25y",
                         age=="26_35" ~ "26 to 35y",
                         age=="36_45" ~ "36 to 45y",
                         age=="46_55" ~ "46 to 55y",
                         age=="56_65" ~ "56 to 65y",
                         age=="65+" ~ "65 & up y")) %>%
  ggplot() +
    geom_hline(aes(yintercept=0), linetype=2) +
    geom_pointrange(aes(y=pct_change_epred_annual, ymin=pct_change_lower_annual, ymax=pct_change_upper_annual, group=acf_period, 
                      x=age,
                      colour = acf_period), size=0.1) +
  scale_y_continuous(labels =percent) +
  facet_grid(.~sex) +
  coord_flip() +
  scale_colour_manual(values = c("#DE0D92", "#4D6CFA")) +
  labs(x="",
       y="Mean annual rate of change in case notification rate (95% UI)\n Before ACF (1950-1956) vs. after ACF (1958-1963)",
       colour="") +
  theme_ggdist() +
  theme(panel.border = element_rect(colour = "grey78", fill=NA))

f3c

```


#### 10.3 Compared to counterfactual

```{r}

counterfact_age_sex <-
      add_epred_draws(object = m_age_sex,
                      newdata = mdata3 %>%
                                    select(year, year2, y_num, age, sex) %>%
                                    mutate(acf_period = "a. pre-acf")) %>%
      filter(year>1957) %>%
      select(year, age, sex, .draw, .epred_counterf = .epred)
  
#Calcuate incidence per draw, then summarise.
  post_change_age_sex <-
      add_epred_draws(object = m_age_sex,
                      newdata = mdata3 %>%
                                    select(year, year2, y_num, age, sex, acf_period)) %>%
      filter(year>1957) %>%
      ungroup() %>%
      select(year, age, sex, .draw, .epred) 
  
  #for the overall period
counterfact_overall_age_sex <-
      add_epred_draws(object = m_age_sex,
                      newdata = mdata3 %>%
                                    select(year, year2, y_num, age, sex) %>%
                                    mutate(acf_period = "a. pre-acf")) %>%
      filter(year>1957) %>%
      select(age, sex, .draw, .epred)  %>%
      group_by(age, sex, .draw) %>%
      summarise(.epred_counterf = sum(.epred)) %>%
      mutate(year = "Overall (1958-1963)")
  
  #Calcuate incidence per draw, then summarise.
  post_change_overall_age_sex <-
      add_epred_draws(object = m_age_sex,
                      newdata = mdata3 %>%
                                    select(year, year2, y_num, age, sex, acf_period)) %>%
      filter(year>1957) %>%
      select(age, sex, .draw, .epred) %>%
      group_by(.draw, age, sex) %>%
      summarise(.epred = sum(.epred)) 
  
  

left_join(counterfact_age_sex, post_change_age_sex) %>%
    mutate(cases_averted = .epred_counterf-.epred,
           pct_change = (.epred - .epred_counterf)/.epred_counterf) %>%
    group_by(year, age, sex) %>%
    mean_qi(cases_averted, pct_change) %>%
    ungroup() %>%
  datatable()

counter_post_overall_age_sex <-
  left_join(counterfact_overall_age_sex, post_change_overall_age_sex) %>%
    mutate(cases_averted = .epred_counterf-.epred,
           pct_change = (.epred - .epred_counterf)/.epred_counterf) %>%
    group_by(age, sex) %>%
    mean_qi(cases_averted, pct_change) %>%
    ungroup() %>%
    mutate(year = "Overall (1958-1963)") 


```


```{r}

age_sex_txt <- counter_post_overall_age_sex %>%
  mutate(across(c(cases_averted:cases_averted.upper), number_format(accuracy = 0.1, big.mark = ","))) %>%
  mutate(across(c(pct_change:pct_change.upper), percent, accuracy=0.1)) %>%
  transmute(year = as.character(year),
            sex = sex,
            age = age,
            cases_averted = glue::glue("{cases_averted}\n({cases_averted.lower} to {cases_averted.upper})"),
            pct_change = glue::glue("{pct_change}\n({pct_change.lower} to {pct_change.upper})"))


age_sex_txt %>% datatable()


```

```{r}

f3d <- counter_post_overall_age_sex %>% 
  mutate(sex = case_when(sex=="M" ~ "Male",
                         sex=="F" ~ "Female")) %>%
  mutate(age = case_when(age=="00_05" ~ "0 to 5y",
                         age=="06_15" ~ "06 to 15y",
                         age=="16_25" ~ "16 to 25y",
                         age=="26_35" ~ "26 to 35y",
                         age=="36_45" ~ "36 to 45y",
                         age=="46_55" ~ "46 to 55y",
                         age=="56_65" ~ "56 to 65y",
                         age=="65+" ~ "65 & up y")) %>%
  ggplot() +
  geom_pointrange(aes(x = age, y=cases_averted, ymin=cases_averted.lower, ymax=cases_averted.upper, colour=sex)) + 
  facet_grid(.~sex) +
  coord_flip() +
  scale_colour_manual(values = c("purple", "darkorange"), name="") +
  scale_y_continuous(labels = comma) +
  labs(x="",
       y="Number (95% UI) of TB cases averted (1958-1963)",
       colour="") +
  theme_ggdist() +
  theme(panel.border = element_rect(colour = "grey78", fill=NA),
        legend.position = "none")

f3d
```




Join together for Figure 2.


```{r}

(f3a + f3b) / (f3c + f3d) + plot_annotation(tag_levels = "A")

ggsave(here("figures/f3.png"), width = 12)


```


### 11. Division-level model

(Very much a work in progress!)

```{r}

mdata4 <- division_inc %>%
  filter(tb_type=="Pulmonary") %>%
  mutate(acf_period = case_when(year %in% c(1950:1956) ~ "a. pre-acf",
                                year %in% c(1957) ~ "b. acf",
                                year %in% c(1958:1963) ~ "c. post-acf")) %>%
  group_by(division) %>%
  mutate(y_num = row_number()) %>%
  ungroup()

```

```{r}
ggplot(data = tibble(x = seq(from = 0, to = 1500, by = 10)),
       aes(x = x, y = dgamma(x, shape = 2, rate = 0.001))) +
  geom_area(color = "transparent", 
            fill = "#DE0D92") +
  scale_x_continuous(NULL) +
  scale_y_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = c(0, 1500)) +
  ggtitle(expression(brms~~gamma(0.5*", "*0.0001)~shape~prior))

winform_prior3 <- c(prior(normal(0, 0.1), class = Intercept),
                  prior(gamma(2, 0.0001), class = shape),
                  prior(normal(0, 0.0001), class = b, coef = "acf_periodb.acf"),
                  prior(normal(0, 0.0001), class = b, coef = "acf_periodc.postMacf"),
                  prior(normal(0, 0.0001), class = b, coef = "y_num"),
                  prior(normal(0, 0.0001), class = b, coef = "y_num:acf_periodb.acf"),
                  prior(normal(0, 0.0001), class = b, coef = "y_num:acf_periodc.postMacf"),
                  prior(cauchy(0,5), class="sd"))


m_pulmonary_division_prior <- brm(
  cases ~ y_num + acf_period + acf_period:y_num + (1 + y_num + acf_period + acf_period:y_num | division ) + offset(log(population_without_inst_ship)),
                  data = mdata4,
                  family = negbinomial(),
                  seed = 1234,
                  chains = 4, cores = 4, 
                  prior = winform_prior3,
                  save_pars = save_pars(all = TRUE),
                  sample_prior = "only",
                  backend = "cmdstanr",
                  warmup = 1000,
  control = list(adapt_delta = 0.9))

conditional_effects(m_pulmonary_division_prior)
plot_counterfactual(model=m_pulmonary_division_prior, model_data=mdata4, population_denominator = population_without_inst_ship, outcome = inc_100k, grouping_var = division, division)


```



```{r}

m_pulmonary_division <- brm(
  cases ~ y_num + acf_period + acf_period:y_num + (1 + y_num + acf_period + acf_period:y_num | division) + offset(log(population_without_inst_ship)),
                  data = mdata4,
                  family = negbinomial(),
                  seed = 1234,
                  chains = 4, cores = 4, 
                  prior = winform_prior3,
                  save_pars = save_pars(all = TRUE),
                  backend = "cmdstanr",
                  warmup = 1000,
  control = list(adapt_delta = 0.9))

summary(m_pulmonary_division)
plot(m_pulmonary_division)
pp_check(m_pulmonary_division, type='ecdf_overlay')

plot_counterfactual(model=m_pulmonary_division, model_data=mdata4, population_denominator = population_without_inst_ship, outcome = inc_100k, grouping_var = division, division)


```

#### 10.2 Summary of impact


```{r}

summarise_change(model_data=mdata4, model = m_pulmonary_division, population_denominator = population_without_inst_ship, grouping_var = division) %>%
  map(datatable)


```



### 12. Counterfactual table

Make a table of counterfactual effects for the manuscript

```{r}

pulmonary_counterfactuals <- tidy_counterfactuals(overall_pulmonary_counterf$counter_post)
pulmonary_counterfactuals_overall <- tidy_counterfactuals_overall(overall_pulmonary_counterf$counter_post_overall)

extrapulmonary_counterfactuals <- tidy_counterfactuals(overall_ep_counterf$counter_post)
extrapulmonary_counterfactuals_overall <- tidy_counterfactuals_overall(overall_ep_counterf$counter_post_overall)

age_sex_counterfactuals_overall <- tidy_counterfactuals_overall(counter_post_overall_age_sex) %>% mutate(model = "Age-sex")

bind_rows(
  bind_rows(pulmonary_counterfactuals, pulmonary_counterfactuals_overall) %>% mutate(model = "Pulmonary TB", sex=NA, age=NA),
  bind_rows(extrapulmonary_counterfactuals, extrapulmonary_counterfactuals_overall) %>% mutate(model = "Extra-pulmonary TB", sex=NA, age=NA),
  age_sex_counterfactuals_overall) %>%
  select(model, year, age, sex, diff_inc, rr_inc, cases_averted, pct_change)



```



##############
#experimental below here
#############


What about a multilevel model with Wards nested within divisions?

```{r}

mdata4 <- ward_inc %>%
  filter(tb_type=="Pulmonary") %>%
  mutate(acf_period = case_when(year %in% c(1950:1956) ~ "a. pre-acf",
                                year %in% c(1957) ~ "b. acf",
                                year %in% c(1958:1963) ~ "c. post-acf")) %>%
  group_by(ward) %>%
  mutate(y_num = row_number()) %>%
  ungroup()



```

```{r}

winform_prior4 <- c(prior(normal(0, 1), class = Intercept),
                  #prior(gamma(1, 0.01), class = shape),
                  prior(normal(0, 1), class = b),
                  prior(cauchy(0,5), class="sd"),
                  prior(lkj(2), class="cor"))
```


```{r}

m_pulmonary_nested <- brm(
  cases ~ y_num + acf_period + acf_period:y_num + (1 + y_num + acf_period + acf_period:y_num | division/ward),
                  data = mdata4,
                  family = negbinomial(),
                  seed = 1234,
                  chains = 4, cores = 4, 
                  prior = winform_prior4,
                  save_pars = save_pars(all = TRUE),
                  backend = "cmdstanr",
                  warmup = 1000)

summary(m_pulmonary_nested)
conditional_effects(m_pulmonary_nested)


```
